### 2021-10-07 paper/header.tex

% **************GENERATED FILE, DO NOT EDIT**************

\title{CompositionalNetworks.jl: a scaling glass-box neural network to learn combinatorial functions}

\author[1]{Jean-François \textsc{Baffier}}
\author[2]{Khalil \textsc{Chrit}}
\author[34]{Florian \textsc{Richoux}}
\author[2]{Pedro \textsc{Patinho}}
\author[2]{Salvador \textsc{Abreu}}
\affil[1]{IIJ, Japan}
\affil[2]{NOVA-LINCS, University of Évora, Portugal}
\affil[3]{AIST, Japan}
\affil[4]{JFLI, CNRS, Japan}

\keywords{Julia Language, Constraint Programming, Local Search, Metaheuristics, Neural Network, Metaprogramming, Scalable Machine Learning, Glass-Box Algorithm}

\hypersetup{
pdftitle = {CompositionalNetworks.jl: a scaling glass-box neural network to learn combinatorial functions},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Jean-François \textsc{Baffier}, Khalil \textsc{Chrit}, Florian \textsc{Richoux}, Pedro \textsc{Patinho}, Salvador \textsc{Abreu}},
pdfkeywords = {Julia Language, Constraint Programming, Local Search, Metaheuristics, Neural Network, Metaprogramming, Scalable Machine Learning, Glass-Box Algorithm},
}